Rethinking Felony Disenfranchisement: The Case of New York City and Miami

### 2019-05-01 markdown/paper.Rmd
change matching procedure from treated if someone entered prison in 2017 to if someone was in prison in 2017
keep writing

##concentrated policing {-}

#Research Design: Spatial Implications of Felony Disenfranchisement in New York City and Miami {-}

##Matching {-}
##Geocoding {-}

>>>>

#Framing for Turnout Effects {-}

Although understanding whether felony disenfranchisement has impacts on turnout among different social groups (namely, racial minorities) is of great importance, understanding the explicitly spatial dimension of the problem is also an important task. Many political decisions bear on the spatial distribution of resources (XX); funding for parks or schools, for instance, can impact all members of a neighborhood, regardless of their race or their income. As such, even a resident who is not directly affected by felony disenfranchisement policies may lose political power thanks to these policies; if many of my neighbors cannot vote, I am joined by fewer voters in advocating for neighborhood-specific investment and services. This is, of course, tied up closely with race; New York City remains highly racially segregated, and therefore disentangling neighborhood effects from race effects becomes difficult. Nevertheless, these effects cross race lines in integrated neighborhoods.

Of course, despite dire over-incarceration in some neighborhoods, the number of incarcerated individuals is relatively low compared to the number of voters. In New York, for instance, `r comma(readRDS("./temp/number_incarcerated.rds"))` individuals were imprisoned in New York State in early 2019, compared with XX actively registered voters. Despite the lower share of residents who are disenfranchised through imprisonment, there is reason to believe the policy impacts more individuals than just those imprisoned. As discussed above, previous research has demonstrated that felony disenfranchisement reduces turnout even among Black voters whose rights are not stripped. This research has found, in particular, that eligible Black voters are less likely to cast a ballot in states where felony disenfranchisement policies are harsher. Understanding the distribution of this effect is important for understanding its political implications. Are voters who live in the same neighborhood as the disenfranchised less likely to cast a ballot because their neighbors have been removed from the community? Or does felony disenfranchisement operate cast its chilling effect on turnout more broadly, crossing even into neighborhoods where voters have not been removed from the community?

This paper begins by examining whether felony disenfranchisement is in fact associated with lower turnout rates in a given neighborhood.^[As discussed below, "neighborhood" is difficult to define. I explore turnout differentials at both the census block group and census tract level.] As discussed above, this depressive effect on turnout is expected to operate through two distinct mechanisms: the actual incarceration and disenfranchisment of convicted individuals, and the social effects on eligible voters identified in previous research. These turnout effects are estimated in various ways: by comparing neighborhoods where voters with a history of participating have been incarcerated with similar neighborhoods where voters have not been incarcerated; by using regression analysis to investigate whether neighborhood arrests are correlated with turnout rates;; and by using regression analysis to investigate the relationship between prison admissions and turnout at the zip-code level.

#Data {-}
##Criminal Justice Data {-}

This paper employs multiple data sets to investigate the extent to which a neighborhood is directly impacted by felony disenfranchisement. The primary data set comes from a freedom of information request filed by the author to obtain individual-level incarceration records for individuals sentenced to incarceration in New York State since 1992. The data includes a host of information, including: first and last name; date of birth; class of offense; incarceration start and end dates; dates of parole; and others. This analysis is limited to individuals incarcerated for felony offenses. Individuals convicted of misdemeanors are not disenfranchised in New York State.

The City of New York also makes the location of all NYPD arrests since 2006 available publicly. This data includes the latitude and longitude of the arrest, the date of the offense, and the class of the alleged offense. This analysis uses only arrests for felony charges.

The third dataset used to estimate the extent to which a neighborhood is directly impacted by felony disenfranchisement policies in New York State comes from another freedom of information request. The individual-level incarceration data referenced above does not include information about the neighborhood an incarcerated person lived in prior to their conviction. Although the State of New York does not make this information available at the individual level, the author has received prison-admission counts by zip-code for the years XX - XX.

##Voter File Data {-}
Most states in the United States are required to maintain files with information on all registered voters in the state. In New York, this information is publicly available from the Board of Elections. It includes information on all registered voters, including: first and last name; date of birth; home address; vote history; and other information. The New York State Voter File also includes information on voters who were previously registered but have since been purged, either because they moved, died, or were incarcerated for a felony offense. Because the voter file includes information on voters who have been purged because of a felony conviction, we are able to reconstruct the vote history of individuals who are no longer registered because of their conviction.

#Method {-}

##Matching {-}
Method.^[Matching on first and last names and dates of birth can result in false positive matches, especially in large states like New York. Following @Meredith2013, I report estimates of the potential impact of these false positives in the Appendix.]
##Geocoding {-}
